### 2019-03-29 2-Chapters/6-Chapter/nonstatbandits.git/NonStatB.tex
Using \begin{framed}...\end{framed} for all algorithms.

% -----------------------------------------------------------------
\section{Introduction}
\label{sec:6:Introduction}

@@

Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation. Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends. Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

@@

In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.

@@

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fix channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.

@@

\paragraph{An interesting interpolation.}
%
The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while the adversarial model can be considered as a special (worst) case, with $\Upsilon_T = T$. However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.

@@

To illustrate the setup presented above, we give here two problems of our benchmark, used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.

% - Pb 1 changes are only on one arm at a time
\textbf{Problem $\bm 1$.}
We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
% (and Figure~\ref{fig:6:Problem_1} in Appendix),
Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
and the optimal arm changes once at $\tau_2^{(1)}=2000$, with a large gap of $0.6$.

@@

% - Pb 2 changes are on all arms at a time
\textbf{Problem $\bm 2$.}
This problem is close to Problem 1, with a minimum optimality gap of $0.1$,
and shown in Figure~\ref{fig:6:Problem_2}.
However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$.
The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$ markers}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$ markers}).

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies are tuned, with a discount factor or a limited memory size, while active strategies are usually tuned with a threshold for the statistical test.

\paragraph{Passively adaptive algorithms.}
%
Their common idea is to adapt a classical policy into forgetting old rewards.
If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes when they happen.
%
The Discounted UCB (D-UCB) algorithm was first introduced in \cite{Kocsis06}, and it is an adaptation of the \UCB{} algorithm, with a discount factor $\gamma\in(0,1)$. It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical mean used for the computation of the \UCB{} indexes.
The regret of D-UCB was proved to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
%
The Sliding-Window UCB (SW-UCB) uses a sliding window of a fixed size $\tau$, to store only the most $\tau$ recent rewards for each arm,
and it was proposed by \cite{Garivier11UCBDiscount}. They proved that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on its regret of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.

Both D-UCB and SW-UCB builds on a stationary policy,but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case,
and achieves a larger regret upper-bound.
We also found in our experiments that it always performed worse that the reference stationary algorithms (\eg, \klUCB), so we do not include it in Section~\ref{sec:6:NumericalExperiments}.
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.
%
For instance, Windowed-Mean Shift from \cite{YuMannor09} is generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting as they consider side observations.
%
A different setting where the quantity of interest is not $\Upsilon_T$ but $V_T$ the total variational budget, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
%
It can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm. Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting. Hence we do no include this algorithm in our experiments.

The idea of using a simple discount factor, as for D-UCB, was recently adapted to a Bayesian policy, with the Discounted Thompson sampling (DTS) algorithm presented by \cite{RajKalyani17}.
Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in their paper.
The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on our Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}). However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
%
In general, we found that passively adaptive approaches can be efficient when their parameters are well tuned, but our experiments show that actively adaptive algorithms perform significantly better.

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley Test and the \UCB{} policy, but no theoretical guarantee was given.
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
On the first hand,
the \CUSUM{} test is rather complicated and parametric, as it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test is similarly complex, and it is also studied in \cite{LiuLeeShroff17},
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments,
and even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
On the other hand, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the different of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound.
As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the time between changes (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
%
\CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
%
\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in their paper).

\CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$. Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$. \CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
\MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$. So it requires the tuning of two parameters, $w$ and $h$. \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.

\paragraph{Expert aggregation.}
%
The second line of research of actively adaptive algorithms uses a Bayesian point of view.
A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
and more recently \cite{Alami17} introduced the Memory Bandit algorithm. It is presented as efficient empirically, but no theoretical guarantee are given, and due to its complexity, we do not include it in our experiments.
The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
Note that the regret bound for these methods are still open-problems,
and they are computationally more costly,
therefore we choose to not include them in our experiments.

\paragraph{Advantages of our approach.}
%
\CUSUMUCB{} and \MUCB{} are both analyzed under some reasonable assumptions on the problem parameters --the means $(\mu_i(t))$-- mostly saying that the breakpoints are sufficiently far away from each other.
%
However, the proposed guarantees only hold for parameters \emph{tuned using some prior knowledge of the means}.
Indeed, while in both cases the threshold $h$ can be set as a function of the horizon $T$ and the number of breakpoints  $\Upsilon_T$ (also needed by previous approaches to obtain the best possible bounds), the parameter $\varepsilon$ for \CUSUM{} and $w$ for \MUCB{} require the knowledge of $\Delta^{\text{change}}$ the smallest magnitude of a change-point.
%
In this chapter, we propose \emph{the first algorithm that does not require this knowledge}, and still attains a $\cO(\sqrt{\Upsilon_T T\log(T)})$ regret.
Moreover we propose the first comparison of the use of local and global restarts within an adaptive algorithm, by studying two variants of our algorithm.
% This study is supported by both theoretical and empirical results presented in this chapter.
%
In others words, if we want to apply our proposal to piece-wise stationary problems such as problem $1$ or $2$ presented above (see Figures~\ref{fig:6:Problem_1}, \ref{fig:6:Problem_2}),
it only needs to know before hand that $T=5000$ and $\Upsilon_T=4$ for these examples, but it does not need to know the amplitude of changes $\max_{i,t} |\mu_i(t) - \mu_i(t+1)|$ nor the optimality gap or a bound on the optimality gap at any time step $\max_{i,j} |\mu_i(t) - \mu_j(t)|$.

>>>>

% -----------------------------------------------------------------
\section{Motivation}
\label{sec:6:Introduction}

@@

% Stochastic bandits were historically introduced as a simple model for clinical trials, where arms correspond to some treatments with unknown efficacy \cite{Thompson33}.
% More recently, MAB models have been proved useful for different applications, like cognitive radio, where arms can model the vacancy of radio channels, as presented in Chapter~\ref{chapter:4}, or parameters of a dynamically configurable radio hardware \cite{Maghsudi16,Bonnefoi17,KerkoucheAlami18}.
% Another application is the design of recommender systems, where arms model the popularity of different items (\eg, news recommendation, \cite{Li10}).
%
For both cognitive radio and recommender systems, the assumption that the arms distribution \emph{do not change over time} may be a big limitation. Indeed, in cognitive radio new devices can enter or leave the network, which impacts the availability of the radio channel they use to communicate; whereas in online recommendation, the popularity of items is also subject to trends. Hence, there has been some interest on how to take those \emph{non-stationary} aspects into account within a multi-armed bandit model.

@@

\TODOL{I need to write a paragraph about using the ``adversarial setting'' for piece-wise stationary. It is very important to state that EXP3.S from \cite{Auer02NonStochastic} (Section~8) achieves $\bigO{\sqrt{K \Upsilon_T T}}$ regret with prior knowledge of $T$ and $\Upsilon_T$.}

% In this chapter, we propose a new algorithm for the piece-wise stationary bandit problem with bounded rewards, called \GLRklUCB. Like previous approaches -- \CUSUM{} \cite{LiuLeeShroff17} and \MUCB{} \cite{CaoZhenKvetonXie18} -- our algorithm relies on combining a standard multi-armed bandit algorithm with a change-point detector. For the bandit component, we propose the use of the \klUCB{} algorithm, that is known to outperform {\UCB}1 used in previous works. For the change-point detector, we propose the Bernoulli Generalized Likelihood Ratio Test (GLRT), for which we provide new non-asymptotic properties that are of independent interest. This choice is particularly appealing because unlike previous approaches, the Bernoulli GLRT is \emph{parameter-free}: it does not need the tuning of a window size ($w$ in \MUCB), or the knowledge of a lower bound on the magnitude of the smallest change ($\varepsilon$ in \CUSUM).

% In this chapter we jointly investigate, both in theory and in practice, two possible combinations of the bandit algorithm with a change-point detector, namely the use of \emph{local restarts} (resetting the history of an arm each time a change-point is detected on that arm) and \emph{global restarts} (resetting the history of \emph{all} arms once a change-point is detected on one of them).
% We provide a regret upper bound scaling in $\cO(\sqrt{\Upsilon_T T \log(T)})$ for both versions of \GLRklUCB{}, matching existing results (when $\Upsilon_T$ is known).
% Our numerical simulations reveal that using local restart leads to better empirical performance, and show that our approach often outperforms existing competitors.

@@

Note than when a breakpoint occurs, we do not assume that all the arms means  change, but that \emph{there exists} an arm whose mean has changed.
Depending on the application, many scenario can be meaningful: changes occurring on all arms simultaneously (due to some exogenous event), or only one or a few arms change at each breakpoint.
For instance, for a cognitive radio application in a IoT-like network, we can imagine that all the objects of one company use a fixed channel or subset of channels, and on a particular day in a particular city, if the company deploys a lot of new objects, then one or some channels see their mean occupancy rates abruptly change.

@@

\paragraph{An interesting interpolation.}
%
The piece-wise stationary bandit model can be viewed as an interpolation between stationary and adversarial models, as the stationary model corresponds to $\Upsilon_T = 0$, while some adversarial models can be considered as a special (worst) case, with $\Upsilon_T = T$ (when considering an adaptive or an oblivious adversary).
However, analyzing an algorithm for the piece-wise stationary bandit model requires to assume a small number of changes, typically $\Upsilon_T = o(\sqrt{T})$.
Note that the adversarial model of \cite{Auer02NonStochastic} is quite powerful, and the authors presented in Section~8 of their paper the EXP3.S algorithm for the piece-wise stationary problem.
\TODOL{Explain here that we should have been more careful regarding this sentence?}

@@

To illustrate the setup presented above, we give here two examples of piece-wise stationary problems.
% problems of our benchmark,
used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.

% - Pb 1 changes are only on one arm at a time
\textbf{Problem $\bm 1$.}
We consider $K=3$ arms changing $\Upsilon_T=4$ times until $T=5000$.
The arm means $(\mu_i(t))_{1\leq i\leq K,1\leq t\leq T}$ are shown in Figure~\ref{fig:6:Problem_1}.
% (and Figure~\ref{fig:6:Problem_1} in Appendix),
Note that changes happen on only one arm (\ie, $C_T=\Upsilon_T=4$),
and the optimal arm changes once at $\tau_2^{(1)}=2000$.
%  with a large gap of $0.6$.

@@

% - Pb 2 changes are on all arms at a time
\textbf{Problem $\bm 2$.}
This problem is close to Problem 1, with a minimum optimality gap of $0.1$ (at any time, the smallest difference between two means is at least $0.1$),
and shown in Figure~\ref{fig:6:Problem_2}.
However, all arms change at every breakpoint (\ie, $C_T=K\Upsilon_T=12$), with identical gaps of $0.1$ for arms $0$ and $1$, and of $0.2$ for arm $2$ (between two breakpoints, the mean change of $+0.1$ for arm $0$ and $-0.1$ for arm $1$ and $-0.2$ for arm $2$).
The first optimal arm decreases at every change (\textcolor{blue}{$2$ with $\nabla$ markers}), and one arm stays the worst (\textcolor{darkgreen}{$1$ with $\diamond$ markers}).

@@

We review previous works that studied the piece-wise stationary bandit model, or variants of this model.
To our knowledge, all the previous works are based on the idea of combining a classical bandit policy, such as Thompson sampling, \UCB{} or EXP3, with a strategy to account for changes in arms' distributions.
Following the vocabulary used in previous works, we make the distinction between passively and actively adaptive strategies.
Active strategies monitors the rewards of each arm by using a change detection algorithm \cite{Basseville93}, and reset the history of pulls and rewards of one or all the arms as soon as a change is detected.
Passive strategies use a (fixed) discount factor or a limited memory size, while most active strategies use a growing memory.
% , while active strategies are usually tuned with a threshold for the statistical test.

\paragraph{Passively adaptive algorithms.}
%
Their common idea is to adapt a classical policy into forgetting old rewards, .
If the forgetting behavior is done efficiently, then the policy can efficiently focus mostly on the most recent rewards, and passively adapt to changes when they happen.
%
The Discounted UCB (D-UCB) algorithm was first introduced in \cite{Kocsis06}, and it is an adaptation of the \UCB{} algorithm, with a discount factor $\gamma\in(0,1)$. It works by decreasing all the past rewards by a multiplicative factor, when receiving a new reward from an arm, so that the recent rewards weight more in the (discounted) empirical mean used for the computation of the \UCB{} indexes.
The regret of D-UCB was proved to be upper-bounded by $\cO(\sqrt{\Upsilon_T T }\log(T))$ in \cite{Garivier11UCBDiscount}, with a tuning $\gamma = 1 - \sqrt{\Upsilon_T/T}/4$, dependent on $\Upsilon_T$.
%
The Sliding-Window UCB (SW-UCB) uses a sliding window of a fixed size $\tau$, to store only the most $\tau$ recent rewards for each arm,
and it was proposed by \cite{Garivier11UCBDiscount}.
They prove that tuning its window-size to $\tau = 2 \sqrt{T\log(T)/\Upsilon_T}$, gives a bound on the regret of SW-UCB of the form $\cO(\sqrt{\Upsilon_T T \log(T)})$.

Both D-UCB and SW-UCB builds on a stationary policy, but for example EXP3.S from \cite{Auer02NonStochastic}
builds on the EXP3 policy, which is designed for the adversarial case.
EXP3.S actually achieves a good regret upper-bound, of the form $\cO(\sqrt{K \Upsilon_T T})$, with no additional prior knowledge except that of $T$ and $\Upsilon_T$.
It constitues a good baseline for our numerical experiments in ,
even if we did find that EXP3.S performs worse that the most of the other approaches based on extensions of stationary bandit algorithms (\eg, SW-UCB).
% , so we do not include it in .
Similarly, previous works showed that older algorithms have no or weaker regret guarantees, and have been proved to be less efficient empirically, or are designed for more specific settings.

The idea of using a simple discount factor, as for D-UCB, was recently adapted to a Bayesian policy, with the Discounted Thompson sampling (DTS) algorithm presented by \cite{RajKalyani17}.
Even if no theoretical guarantee was given, it can be empirically very efficient, but we found that DTS is not robust in the choice of if its discount factor $\gamma\in(0,1)$, contrarily to what was highlighted in the paper.
The DTS algorithm can perform well in practice, for instance with $\gamma=0.75$ on Problems $1$ and $2$ (see Section~\ref{sec:6:NumericalExperiments}). However, no theoretical guarantees are given for this strategy, and our experiments did not really confirm the robustness to $\gamma$.
%
In general, we found that passively adaptive approaches can be efficient when their parameters are well tuned, but our experiments show that actively adaptive algorithms perform significantly better.

Finally, XXX use two discount factors at the same time, see the NEWMA algorithm from \cite{Keriven2018}.

\TODOL{Talk about NEWMA and its possible application to generalize/improve D-UCB, \cite{Keriven2018}.}

\paragraph{Actively adaptive algorithms.}
%
We distinguish two families,
the first line of research uses frequentist change-point detectors \cite{Basseville93}, combined with stationary policies, usually index policies like \UCB.
When using an efficient CD algorithm with an efficient index policy, these approaches usually perform more efficiently than the passively algorithms.
%
The Adapt-EVE algorithm from \cite{Hartland06} uses a Page-Hikley (PH) Test and the \UCB{} policy, but no theoretical guarantee was given.
%
The Windowed-Mean Shift algorithm from \cite{YuMannor09} is more generic and combines any efficient bandit policy with a CD test based on a sliding window, but there approach is not applicable to the bandit setting of interest in this chapter, as they consider side observations.
%
The EXP3.R algorithm from \cite{Allesiardo15,Allesiardo17} combines a CD algorithm with EXP3, and the history of all arms are reset as soon as a suboptimal arm is detecting to become optimal.
A regret bound of $\cO(\Upsilon_T \sqrt{T \log(T)})$ was proved, even when $\Upsilon_T$ is known.

Two recent and related works use
the two-sided \CUSUM{} CD algorithm for \CUSUMUCB{} in \cite{LiuLeeShroff17}
or a specific and simpler CD algorithm for the Monitored UCB (\MUCB) algorithm introduced in \cite{CaoZhenKvetonXie18}.
%
The \CUSUM{} test is rather complicated and parametric: it uses the first $M$ samples (for a fixed $M$) from one arm to compute an average $\hat{u_0}$, and then builds two random walks, using the remaining observations for this arm. A change is detected when either random walks cross a threshold $h$.
It requires the tuning of three parameters, $M$ and $h$ as well as a drift correction parameter $\varepsilon\in(0,1)$.
The PH test, also studied in \cite{LiuLeeShroff17}, is similarly complex,
and we found that \PHT-\UCB{} perform very similarly to \CUSUMUCB{} in our experiments.
Even if it has the advantage of not requiring a parameter $M$,
it has no theoretical guarantee, so we do not include \PHT-\UCB{} in the experiments presented below.
%
In comparison, \MUCB{} uses a much simpler test, based on the $w$ most recent observations on one arm (for a fixed and even number $w$), and compares with a fixed threshold $h$ the difference of the sum of rewards for the first half and the second half. It is numerically much simpler, and has the advantage of using only a bounded memory (of the order $\bigO{K w}$ for $K$ arms).
%
Both approaches also introduced a mechanism to force a uniform exploration of each arm, parametered by $\alpha\in(0,1)$.
% \CUSUMUCB{} uses a randomized exploration, by sampling arm $i$ with probability $\alpha/K$ at each time step,
% and \MUCB{} uses a more deterministic scheme, to force exploring the $K$ arms (when the time since the last restart, $t - \tau$, is found to be in $\{1,\dots,K\}$ modulo $\left\lceil 1/\alpha \right\rceil$).
% Both solutions ensure that all arms are sampled enough on each stationary sequence, in order for the CD algorithm to have enough \iid{} samples to detect changes.

When all their parameters are tuned correctly, both approaches are proved to achieve a good regret upper-bound, of order $\cO(K \sqrt{\Upsilon_T T \log(K T)})$.
As both results are valid under different assumptions, we consider the two results to be the current state-of-the-art.
As for previous works, tuning their parameters requires to know both the horizon $T$ and the number of breakpoints $\Upsilon_T$, but most importantly it requires a prior knowledge of the problem difficulty, by assuming to know a lower bound on both the length of stationary segments (\eg, $L$ for \MUCB) and on the changes on the means of arms (\eg, $\varepsilon$ for \CUSUM).
%
\CUSUMUCB{} achieves a better regret upper bound, of the order of $\cO(\sqrt{\Upsilon_T T \log(T / \Upsilon_T)})$, but \emph{only} for Bernoulli distributions,
and when its $4$ parameters ($\alpha,\varepsilon,M,h$) are tuned based on problem-dependent knowledge.
%
\MUCB{} achieves a regret bounded by $\cO(\sqrt{\Upsilon_T T \log(T)})$, for bounded distributions, and when its $4$ parameters ($\alpha,w,b,\gamma$) are also tuned based on a problem-dependent knowledge of $\tilde{\delta}$ and $L$ (see Assumption~1 and Remark~1 in \cite{CaoZhenKvetonXie18}).

% \CUSUMUCB{} is based on a rather complicated two-sided \CUSUM{} test, that uses the first $M$ samples from one arm to compute an initial average, and then detects whether a drift of size larger than $\varepsilon$ occurred from this value by checking whether a random walk based on the remaining observations crosses a threshold $h$. Thus it requires the tuning of three parameters, $M$, $\varepsilon$ and $h$.
On the one hand, \CUSUMUCB{} performs \emph{local restarts} using this test, to reset the history of \emph{one arm} for which the test detects a change.
% \MUCB{} uses a much simpler test, based on the $w$ most recent observations from an arm: a change is detected if the absolute difference between the empirical means of the first and second halves of those $w$ observations exceeds a threshold $h$. So it requires the tuning of two parameters, $w$ and $h$.
On the other hand, \MUCB{} performs \emph{global restarts} using this test, to reset the history of \emph{all arms} whenever the test detects a change on one of the arms.
Compared to \CUSUMUCB{}, note that \MUCB{} is numerically much simpler as it only uses a bounded memory, of order $\bigO{K w}$ for $K$ arms.

\paragraph{Expert aggregation.}
%
Another of research on actively adaptive algorithms uses a Bayesian point of view.
A Bayesian CD algorithm is combined with Thompson Sampling in \cite{MellorShapiro13},
and more recently \cite{Alami17} introduced the Memory Bandit algorithm.
It is presented as efficient empirically, but no theoretical guarantee are given for these two works, and due to its complexity, we do not include it in our experiments.
The idea behind Memory Bandit is to use an expert aggregation algorithm, like EXP4 from \cite{Auer02}, modified to efficiently aggregate an growing number of experts from techniques presented in \cite{Mourtada17}.
At each time step, a new expert is introduced, and experts corresponds to different Thompson sampling algorithms, each using a different history of pulls and rewards. Intuitively, after a change-point the newest experts will soon become the most efficient, as they are learning by using rewards drawn from the new distribution(s).
%
Note that the regret bound for these methods are still open-problems,
and they are computationally more costly,
therefore we choose to not include them in our experiments.

\paragraph{Slowly-varying model.}

A different setting where the quantity of interest is not $\Upsilon_T$ but the total variational budget $V_T$, was introduced by \cite{Besbes14stochastic}, for which they proposed the RExp3 algorithm.
The total variation budget $V_T$ is defined as $\sum_{t=1}^{T-1} \sum_{i=1}^K D(\nu_i(t), \nu_i(t+1))$, for a certain measure of difference $D$, which measures how much the two distributions of any arm $i$ have changed from time $t$ to time $t+1$.
Two examples can be the total variation distance $\|\cdot\|_{TV}$ or the Kullback-Leibler divergence.
%
The RExp3 algorithm can also be qualified as passively adaptive: it is based on (non-adaptive) restarts of the EXP3 algorithm. Note that this algorithm is introduced for a different setting, where the quantity of interest is not $\Upsilon_T$ but a quantity $V_T$ called the total variational budget (satisfying $\Delta^{\text{change}} \Upsilon_T \leq V_T \leq \Upsilon_T$ with $\Delta^{\text{change}}$ the minimum magnitude of a change-point).
A regret bound of the order of $\cO(V_T^{1/3} T^{2/3})$ is proved, which is weaker than existing results in our setting.

Three recent works proposed algorithms that achieve a small regret under the two models.
For simple bandits, \cite{WeiSrivastava18Abruptly} proposed the SW-UCB\# and LM-DSEE algorithms FIXME.
For contextual bandits, \cite{Luo18} and \cite{ChenLeeLuoWei2019} FIXME.
\TODOL{Parler de ces travaux là, qui sont très forts mais pas non plus encore parfaits.}

This setting is quite different from the setting we are interested by in this chapter, and even if the most recent works on slowly-varying MAB models are very interesting, we preferred to focus on the piece-wise stationary (or abruptly changing) model,
hence we do no include any of these algorithms in the experiments presented in Section~\ref{sec:6:NumericalExperiments}.

\paragraph{Another point-of-view on prior knowledge of $\Upsilon_T$.}
%
In this chapter, like in our two main inspirational works \cite{CaoZhenKvetonXie18,LiuLeeShroff17},
we assume that the algorithm know in advance the number of break-points $\Upsilon_T$.
Another interesting point-of-view is the one presented by \cite{WeiSrivastava18Abruptly,WeiSrivastava18Distributed},
where the authors assume that there exists a number $\nu\in(0,1)$ for which $\Upsilon_T = \smallO{T^{\nu}}$,
and they also assume that any algorithm tackling such problems can know in advance the value of $\nu$ (or an upper-bound on $\nu$), and tweak its parameters from this value $\nu$, but cannot know the exact value of $\Upsilon_T$.
%
This interpretation is interesting too, but due to time and space constraints, we preferred to not explore it further.
Empirically, we found that the SW-UCB\# algorithm proposed in \cite{WeiSrivastava18Abruptly} performs comparably to SW-UCB, and thus we did not include it in the experiments presented below in Section~\ref{sec:6:NumericalExperiments}.
